Capstone Movielens Report

### 2025-06-04 reports/capstone-movielens.site/02.8-1.MF.UMGYDE-model.Rmd

$$
\begin{array}{cc} min \\
\tiny{P,Q} \end{array} \sum_{(u,v)\in R}[f(\mathbf{p}_u, \mathbf{q}_v; \mathbf{r}_{u,v}) + \mu_P\parallel\mathbf{p}_u\parallel_1 + \mu_Q\parallel\mathbf{q}_v\parallel_1 + \frac{\lambda_P}2\parallel\mathbf{p}_u\parallel^2_2 + \frac{\lambda_Q}2\parallel\mathbf{q}_v\parallel^2_2]
$$

@@

>>>>

$$
\begin{array}{c} { min} \\
{\scriptstyle P,Q} \end{array} \sum_{(u,v)\in R}[f(\mathbf{p}_u, \mathbf{q}_v; \mathbf{r}_{u,v}) + \mu_P\parallel\mathbf{p}_u\parallel_1 + \mu_Q\parallel\mathbf{q}_v\parallel_1 + \frac{\lambda_P}2\parallel\mathbf{p}_u\parallel^2_2 + \frac{\lambda_Q}2\parallel\mathbf{q}_v\parallel^2_2]
$$

@@

### MF: Performing Operation

::: {.noteblock data-latex=""}
The complete source code shown in this section is available in the [Perform the Matrix Factorization & Final Test](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/capstone-movielens.main.R#L2567) section of the [capstone-movielens.main.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/capstone-movielens.main.R#L2567) script on _GitHub_.
:::

We assume that the _Matrix Factorization_ method is going to be the last one to provide sufficient results, so we will use the entire `edx` dataset as a _Training Set_ and `final_holdout_test` as a _Test Set_ for our final model, skipping the intermediate _train/test_ operations on the split _train/validation_ datasets.

#### MF: Getting Residuals From the UMGYDE Model Prediction Values
\

First, we get residuals from the _UMGYDE Model_ prediction values using the [mf.residual.dataframe](#func.mf.residual.dataframe) function described above:
```{r, eval=FALSE}
  mf.edx.residual <- mf.residual.dataframe(edx)
```
```{r}
  str(mf.edx.residual)
```
#### MF: Transforming Input Data to Be Compatible With the `recosystem` Package
\

Now, let's convert the residuals data and `final_holdout_test` dataset to the input data compatible with the [recosystem](https://cran.r-project.org/web/packages/recosystem/index.html) package to process:
```{r, eval=FALSE}
  set.seed(5430)
  mf.edx.residual.reco <- with(mf.edx.residual, 
                               data_memory(user_index = userId, 
                                           item_index = movieId,
                                           rating = rsdl))

  final_holdout_test.reco <- with(final_holdout_test, 
                          data_memory(user_index = userId, 
                                      item_index = movieId, 
                                      rating = rating))
```
```{r}
str(mf.edx.residual.reco)
str(final_holdout_test.reco)
```

#### MF: Creating and Tuning the `Reco` Object
\

We then create a `reco` object of the `Reco` class and tune it using the `reco$tune` method, passing to it as a `train_data` parameter residual data prepared in the previous step:
```{r, eval=FALSE}
  reco <- Reco()

  reco.tuned <- reco$tune(mf.edx.residual.reco, opts = list(dim = c(10, 20, 30),
                                                lrate    = c(0.1, 0.2),
                                                nthread  = 4,
                                                niter    = 10,
                                                verbose  = TRUE))
```
```{r}
str(reco)
str(reco.tuned)
```

#### MF: Final Training
\

Finally, we train the model for the last time using the `reco$train` method, passing to it as a `train_data` parameter the same residual data we used for the tuning:
```{r, eval=FALSE}
  reco$train(mf.edx.residual.reco, opts = c(reco.tuned$min,
                                       niter = 20, 
                                       nthread = 4)) 
```

#### MF: Final Holdout Test
\

From now on, we will work with the `final_holdout_test` dataset for the final testing of our model, obtaining final predictions, and computing the _Root Mean Squared Error_.

First, we get the final predictions combining the predicted residuals from the `reco` objects and predicted values we lastly get from the _UMGYDE Model_ as described in the [UMGYDE Model: Final Holdout Test (Preliminary Assessment)] section:
```{r, eval=FALSE}
  mf.reco.residual <- reco$predict(final_holdout_test.reco, out_memory())

  mf.predicted_ratings <- 
    clamp(final.UMGYDE.predicted$predicted + mf.reco.residual)
```
```{r}
str(mf.reco.residual)
str(mf.predicted_ratings)
```
Next, we compute the _RMSE_ using the [rmse2](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/common-helper.functions.R#L28) custom helper function, already mentioned above in the [UMGYDE Model: Final Holdout Test (Preliminary Assessment)] section:
```{r, eval=FALSE}
final_holdout_test.RMSE <- rmse2(final_holdout_test$rating,
                                     mf.predicted_ratings)
```

::: {.noteblock data-latex=""}
The function [rmse2](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/common-helper.functions.R#L28) is defined in the  [(R)MSE-related functions](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/common-helper.functions.R#L18) section of the [common-helper.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/common-helper.functions.R#L28) script on _GitHub_.
:::

And add the result to our _Result Table_:
```{r, eval=FALSE}
final.MF.RMSEs.ResultTibble <- final.RMSEs.ResultTibble.UMGYDE.rglr.tuned |> 
  RMSEs.AddRow("MF (Final Test)", 
               final_holdout_test.RMSE,
               comment = "Matrix Factorization of the Best Model Residuals, Final Holdout Test")
```
```{r}
RMSE_kable(final.MF.RMSEs.ResultTibble)
```

::: {.noteblock data-latex=""}
Finally, we have reached our ultimate goal, since the _Root Mean Squared Error_ we obtained has achieved the required _Project Objective_.
:::
